Radiocarbon measurement of CO~2~ from carbonate minerals using a hybrid gas ion source with an open split interface

### 2022-04-13 doc/AMS-15/AMS-15_paper/AMS_HGIS_els/AMS_HGIS_els.Rmd
Format figure axes, add supplementary table + data, lots of minor edits, improve references.

---
title: "Radiocarbon measurement of CO~2~ from carbonate minerals using a hybrid gas ion source with an open split interface"
author:
  - name: Brett Longworth
    email: blongworth@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
    correspondingauthor: true
  - name: Josh Burton
    email: jburton@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
  - name: Simon Pendleton
    email: simon.pendleton@plymouth.edu
    affiliation:  Woods Hole Oceanographic Institution,Plymouth State University
  - name: Sydney Moser
    email: smose007@uottawa.ca
    affiliation: University of Ottawa
  - name: Mark Roberts
    email: mroberts@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
  - name: Mark Kurz
    email: mkurz@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
address:
  - code: Woods Hole Oceanographic Institution
    address: Department of Geology and Geophysics, 360 Woods Hole Rd., Woods Hole, MA, 02540
  - code: Plymouth State University
    address: Environmental Science and Policy, 17 High St., Plymouth, MA, 03264
  - code: University of Ottawa
    address: Geography, Environment and Geomatics, 550 Cumberland St., Ottawa, ON, Canada, K1N 6N5
abstract: |
  We have developed a method for measuring radiocarbon in carbonate minerals as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS). The method uses helium as a carrier gas to displace CO~2~ from sample vials to an open split, where a glass capillary samples the mixture for delivery directly to the HGIS. This method skips the gas transfer and quantification steps used in a closed inlet HGIS system, simplifying sample measurement. Samples larger than 8 mg carbonate can be measured. Results from measurements of consensus standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples from a marine core (F^14^C = 0.4 - 1.15) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the 1$\sigma$ uncertainty is about 1 %. We discuss advantages and disadvantages of continuous flow sample introduction, and the effect of reduced precision on calibrated age-depth models produced using gas-source data.
keywords: 
  - radiocarbon
  - carbonate
  - hybrid gas ion source
journal: "Nuclear Instruments and Methods in Physics Research Section B"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: ../HGIS_paper.bib
linenumbers: false
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
csl: elsevier-with-titles.csl
output: 
  rticles::elsevier_article:
    latex_engine: xelatex
    keep_tex: true
    citation_package: default
---

@@

```{r, message = FALSE}
library(tidyverse)
library(hgis)
library(here)
#library(errors)

#options(errors.notation = "plus-minus")

@@

Radiocarbon Accelerator Mass Spectrometry (AMS) is a sensitive method for determining the age, provenance, and composition of carbon-containing materials. A primary disadvantage of the method is cost, which can limit the number of samples analyzed. This forces researchers to carefully select samples for analysis and may limit repeat measurements for error determination. One driver of the cost of AMS analysis is the time and effort required for sample preparation, including reduction of the sample to graphite. Analysis of CO~2~ directly via a gas-accepting AMS ion source eliminates the cost and effort of graphitization and returns results more rapidly. The NOSAMS USAMS system at the National Oceanographic AMS lab (NOSAMS) is based on a 3MV Tandetron accelerator [@Longworth2015] combined with a National Electrostatics Corporation (NEC) 134-sample gas-capable MCSNICS source [@Xu2007] . We present here a system for semi-automated analysis of carbonate minerals by acid hydrolysis in vials and introduction to the ion source in a flow of helium via an open split.

@@

The open split interface (Figure \ref{fig1}) supplies a helium - CO~2~ mixture to a split where a capillary transfers gas to the ion source and the gas not taken by the capillary is vented. Flowing helium is used to displace sample CO~2~ from the sample vial to the delivery capillary in the split. The open-split interface is constructed of simple, off-the-shelf parts. The core of the design is a dual needle and split housed in a Valco tee fitting. Displacement helium flows in through the sidearm of the tee and into the sample vial through 1/16 in. stainless steel tubing, which forms the outer flow path of the concentric dual needle.

@@

The HGIS measurements were carried out on a modified 3MV Tandetron AMS system [@Longworth2015]. Measurement of a sample proceeded as follows: a clean sputter target cathode was rotated into position and the target was pre-cleaned by sputtering for 5 minutes or until cathode ^12^C^-^ current fell below 100 nA. The delivery needle was then transferred to the sample vial, which started gas delivery to the split and source. Ion current was allowed to stabilize for 30 s, at which point the data acquisition was started. Data was acquired in 10 blocks of 90 s each, for 15 min total acquisition time. The displacement helium was allowed to flow constantly during cathode cleaning and analysis, and the sample needle was inserted into a helium-filled vial between samples. Normalizing standards, blanks, and secondary standards were measured identically to unknowns. Total time for a measurement was 20-25 min including target pre-cleaning and changing samples. Graphite samples were prepared at NOSAMS and measured on the CFAMS system using standard procedures [@gagnon_nosams_2000; @roberts_high-performance_2010].

@@

Initial work with the NEC source established optimal source parameters for consistent source performance. We used tank CO~2~ connected through a regulator and glass capillary to the ion source to measure source performance over a range of CO~2~ flows. Best ^12^C^-^ ion current for the source was 10 $\mu$A, or roughly 10% of typical graphite target current. Peak current was produced at a CO~2~ flow rate of 7.5 $\mu$l min^-1^ CO~2~, but the source produced near optimal currents with flows from 7.5-25 $\mu$l min^-1^ CO~2~. This is comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance [@Salazar2013; @Fahrni2013; @broek_optimization_nodate]. Typical currents for this study were between 4-9 $\mu$A and average current for all samples was `r curmean` ± `r cursd` $\mu$A. Several studies have shown best currents at lower CO~2~ flow rates [@Xu2007; @Ruff2010].

@@

```{r}
# Fix to use current form of processed data
blanks <- carb %>% 
  filter(rec_num %in% c(2138, 83028),
         he12C > 2E-6,
         sample_type == "B")
blankmean <- mean(blanks$norm_ratio) %>% format(digits = 3)
blanksd <-  sd(blanks$norm_ratio) %>% format(digits = 2)

@@

The HGIS method produced higher $^{14/12}C$ ratios for radiocarbon dead samples than the graphite method. The average uncorrected $F^{14}C$ for radiocarbon-free reference materials IAEA C1 marble and TIRI F spar calcite was `r blankmean` ± `r blanksd` (1$\sigma$ uncertainty, N=`r nblank`), compared to XXX for graphite. We investigated the sources of this method blank as a current-dependent blank and as a current-independent blank, and implemented a blank correction for the method.

HGIS cathodes produce measurable ion current, even with no gas flow. This current was initially \> 10 $\mu$A and dropped exponentially, reaching a baseline of ~100 nA after 5-10 minutes. For the range of currents in this study (3-9 $\mu$A), this baseline current represents a 1-6% contribution to the total measured sample current at the F^14^C of the target current baseline. This can be interpreted as a constant current dependent blank, analogous to a mass-dependent "constant contaminant" blank and may be modeled with the same tools [@haghipour_compound-specific_2019; @hanke_comprehensive_2017; @roberts_14c_2019; @sun_14_2020]. To assess blank contribution and ratio dependence on sample current, vials of live and dead standard gas were measured as the displacement helium flow diluted the samples, producing diminishing current over time. We used these data to develop a current-balance model following methods used to assess mass-dependent or constant contamination blanks in graphite targets [@shah_walter_ultra-small_2015; @sun_14_2020]. The constant contamination blank was also assessed directly by measuring pure helium in vials using the same method. Results from both tests are shown in Figure 3a. The inverse of the sample current is used to linearize the data, allowing assessment with a simple linear model. Fits from modern and dead dilution tests should trend towards the fraction modern of the blank, crossing at the inverse of blank current. Based on this assessment, we estimate the F^14^C of the blank as 0.43 with a current of 110 nA.

```{r fig3, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig3}Assessment of blanks. (a) Current dependent blank model and data. $F^{14}C$ and current of measurements from dilution tests of modern and dead tank standards are used to generate linear fits which should cross at the $F^{14}C$ and current of the constant current blank. Direct determination of $F^{14}C$ and current of the blank using clean targets and helium without $CO_2$ are also shown. (b) Fraction modern of large IAEA C1 and TIRI-F carbonate blank samples compared to sample current. Horizontal and vertical bars represent measurement uncertainty in current and ratio, respectively.", echo = FALSE}
knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig3_blanks.pdf"))
```

@@

options(digits = 3)
options(scipen = 4)
options(decimals = 4)
options(errors.notation = "plus-minus")

@@

Reference materials analyzed with the hybrid gas method were in good agreement with consensus values. Mean difference from consensus for all measurements was F^14^C `r format(cons_sum$fm_diff_mean, digits = 1, nsmall = 4)` shows that the hybrid method has no bias in aggregate. The standard deviation of the difference in F^14^C was `r format(cons_sum$fm_diff_sd, nsmall = 4)`, which may be taken as an estimate of the 1$\sigma$ precision of the method. The mean of propagated measurement errors for these samples was `r format(cons_sum$sig_fm_corr_mean, nsmall = 4)` F^14^C for these samples, which is in good agreement with the method precision above. The method error is about three times larger than the 2--3 ‰ precision of graphite measurements, but is within the range of method precision of other gas sources.

```{r tab1, message = FALSE}
# Summary table
cons_df %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  ungroup() %>% 
  arrange(fm_consensus) %>% 
  mutate (fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd),
          Name = case_when(Name == "TIRI-I" ~ "TIRI I",
                           Name == "C-2" ~ "IAEA C2",
                           Name == "NOSAMS2" ~ "NOSAMS 2")) %>% 
  select(`Sample Name` = Name, 
         `Reference F^14^C` = fm_consensus,
         `HGIS F^14^C ± stdev` = fm_corr_err,
         `Mean per-sample error` = sig_fm_corr_mean,
         N) %>% 
  knitr::kable(digits = 3,
               caption = "\\label{tab1}Reference material precision and accuracy. Reference 
               values are community consensus values for TIRI I and C2, and the 
               NOSAMS internal graphite value for NOSAMS2. The mean of per-sample 
               error estimates is included for comparison to the reproducibilty 
               of the samples.")

@@

```{r fig4, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig4}Agreement of standards with graphite consensus values. Difference of HGIS measurements from consensus for individual samples are plotted against their expected value. Expected values are spread to avoid overlapping points. Error bars represent 1 $\\sigma$ sample uncertainty. The black line is the mean difference from expected value for all measurements, and the blue lines are one standard deviation of all measurements.", echo = FALSE}
knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig4_consensus_samples.pdf"))
```

@@

```{r fig5, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig5}Comparison of Haiti core samples measured as graphite and via HGIS. (a) Difference between HGIS and graphite measurements of the same materials. Difference between HGIS and graphite measurements is shown on the vertical axis and plotted against the mean of the two measurements. Error bars are 1 $\\sigma$ measurement uncertainty for each sample. Graphite measurements are included to allow comparison of uncertainty. Age-depth models produced using HGIS (b) and graphite measurements (c). Grey subplots show probability distribution of calibrated ages. (d) Combined age-depth prediction and 95\\% confidence intervals for HGIS and graphte. (e) Offset of HGIS and graphite values from the mean of the models with confidence intervals.", echo = FALSE}
knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig5_core_carbonates.pdf"))
```

@@

Age models were developed for both the HGIS measurements (Figure 5b) and the traditional graphite measurements (Figure 5c) of the core carbonates. We used the Bchron Baesian radiochronology package [@haslett_simple_2008] to develop these models. We applied a correction for the regional marine reservoir effect with a $\Delta$ R of -216 ± 92 years based on the Oxcal model [@ramsey_bayesian_2009] with 10 reservoir points. Samples with pre-bomb radiocarbon ages were calibrated using the Marine20 curve @heaton_marine20marine_2020. The NH zone 2 curve from Hua et. al @hua_atmospheric_2013 was used for calibration of postmodern dates, but these measurements were not included in the age model due to uncertainties in reservoir correction. The age-depth curves produced by the two datasets agree within the uncertainty of the chronologies, with the largest differences being in the uncertainty of the individual dates.

@@

We developed a method for the analysis of carbonates using an open-split interface to a hybrid sputter gas ion source. This system produces results that agree with graphite measurements within the approximately 1% method error from 8-20 mg of carbonate. Sample preparation takes about 10 min. of person time and 6 h of automated processing and reaction time, and 20 min. of measurement time per sample. The simplicity of of this approach can potentially reduce sample cost by eliminating graphitization, and may improve reliability and reduce maintenance compared to more complex sample inlet designs.

>>>>

---
title: "Radiocarbon measurement of CO~2~ from carbonate minerals using a hybrid gas ion source with an open split interface"
author:
  - name: Brett E. Longworth
    email: blongworth@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
    correspondingauthor: true
  - name: Josh R. Burton
    email: jburton@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
  - name: Simon Pendleton
    email: simon.pendleton@plymouth.edu
    affiliation:  Woods Hole Oceanographic Institution,Plymouth State University
  - name: Sydney D. Moser
    email: smose007@uottawa.ca
    affiliation: University of Ottawa
  - name: Mark L. Roberts
    email: mroberts@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
  - name: Mark D. Kurz
    email: mkurz@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
address:
  - code: Woods Hole Oceanographic Institution
    address: Department of Geology and Geophysics, 360 Woods Hole Rd., Woods Hole, MA, 02540
  - code: Plymouth State University
    address: Environmental Science and Policy, 17 High St., Plymouth, MA, 03264
  - code: University of Ottawa
    address: Geography, Environment and Geomatics, 550 Cumberland St., Ottawa, ON, Canada, K1N 6N5
abstract: |
  We have developed a method for measuring radiocarbon in carbonate minerals as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS). The method uses helium as a carrier gas to displace CO~2~ from sample vials to an open split, where a glass capillary samples the mixture for delivery directly to the HGIS. This method skips the gas transfer and quantification steps used in a closed inlet HGIS system, simplifying sample measurement. Samples larger than 8 mg carbonate can be measured. Results from measurements of consensus standards (TIRI I, IAEA C2, and an internal modern shell standard), and samples from a marine core (F^14^C = 0.4--1.15) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the 1$\sigma$ uncertainty is about 1 %. We discuss advantages and disadvantages of continuous flow sample introduction, and the effect of reduced precision on calibrated age-depth models produced using gas-source data.
keywords: 
  - radiocarbon
  - carbonate
  - hybrid gas ion source
journal: "Nuclear Instruments and Methods in Physics Research Section B"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: ../HGIS_paper.bib
linenumbers: false
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
csl: elsevier-with-titles.csl
output: 
  rticles::elsevier_article:
    latex_engine: xelatex
    keep_tex: true
    citation_package: default
---

@@

```{r, message = FALSE}
library(tidyverse)
library(hgis)
library(here)
library(errors)

options(errors.notation = "plus-minus")

@@

Radiocarbon Accelerator Mass Spectrometry (AMS) is a sensitive method for determining the age, provenance, and composition of carbon-containing materials. A primary disadvantage of the method is cost, which can limit the number of samples analyzed. This forces researchers to carefully select samples for analysis and may limit repeat measurements for error determination. One driver of the cost of AMS analysis is the time and effort required for sample preparation, including reduction of the sample to graphite. Analysis of CO~2~ directly via a gas-accepting AMS ion source eliminates the cost and effort of graphitization and returns results more rapidly. The NOSAMS USAMS system at the National Oceanographic AMS lab (NOSAMS) is based on a 3MV Tandetron accelerator [@Longworth2015] combined with a National Electrostatics Corporation (NEC) 134-sample gas-capable MCSNICS source. The hybrid gas capability of this source was developed as a joint project between NEC and the Scottish Universities Environmental Research Centre (SUERC) [@Xu2007], and further improved by work at Lawrence Livermore National Laboratory (LLNL) [@Salazar2013]. We present here a system for semi-automated analysis of carbonate minerals by acid hydrolysis in vials and introduction to the ion source in a flow of helium via an open split.

@@

The open split interface (Figure \ref{fig1}) supplies a CO~2~ - helium mixture to a split where a capillary transfers gas to the ion source and the gas not taken by the capillary is vented. Flowing helium is used to displace sample CO~2~ from the sample vial to the delivery capillary in the split. The open-split interface is constructed of simple, off-the-shelf parts. The core of the design is a dual needle and split housed in a Valco tee fitting. Displacement helium flows in through the sidearm of the tee and into the sample vial through 1/16 in. stainless steel tubing, which forms the outer flow path of the concentric dual needle.

@@

The HGIS measurements were carried out on a modified 3MV Tandetron AMS system [@Longworth2015]. Measurement of a sample proceeded as follows: a clean sputter target cathode was rotated into position and the target was pre-cleaned by sputtering for 5 minutes or until cathode ^12^C^-^ current fell below 100 nA. The delivery needle was then transferred to the sample vial, which started gas delivery to the split and source. Ion current was allowed to stabilize for 30 s, at which point the data acquisition was started. Data was acquired in 10 blocks of 90 s each, for 15 min total acquisition time. The displacement helium was allowed to flow constantly during cathode cleaning and analysis, and the sample needle was inserted into a helium-filled vial between samples. Normalizing standards, blanks, and secondary standards were measured identically to unknowns. One target is used per gas sample. Total time for a measurement was 20-25 min including target pre-cleaning and changing samples. Graphite samples were prepared at NOSAMS and measured on the CFAMS system using standard procedures [@gagnon_nosams_2000; @roberts_high-performance_2010].

@@

Initial work with the NEC source established optimal source parameters for consistent source performance. We used tank CO~2~ connected through a regulator and glass capillary to the ion source to measure source performance over a range of CO~2~ flows. Best ^12^C^-^ ion current for the source was 10 $\mu$A, or roughly 10% of typical graphite target current. Peak current was produced at a CO~2~ flow rate of 7.5 $\mu$l min^-1^ CO~2~, but the source produced near optimal currents with flows from 7.5--25 $\mu$l min^-1^ CO~2~. This is comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance [@Salazar2013; @Fahrni2013; @broek_optimization_nodate]. Typical currents for this study were between 4-9 $\mu$A and average current for all samples was `r curmean` ± `r cursd` $\mu$A. Several studies have shown best currents at lower CO~2~ flow rates [@Xu2007; @Ruff2010].

@@

```{r}
# Fix to use current form of processed data
blanks <- carb %>% 
  filter(rec_num %in% c(2138, 83028),
         he12C > 2E-6,
         sample_type == "B")
blankmean <- mean(blanks$norm_ratio) %>% format(digits = 2)
blanksd <-  sd(blanks$norm_ratio) %>% format(digits = 1)

@@

The HGIS method produced higher $^{14/12}C$ ratios for radiocarbon dead samples than the graphite method. The average uncorrected $F^{14}C$ for radiocarbon-free reference materials IAEA C1 marble and TIRI F spar calcite was `r blankmean` ± `r blanksd` (1$\sigma$ uncertainty, N=`r nblank`), compared to 0.003 ± 0.001 for graphite. We investigated the sources of this method blank as a current-dependent blank and as a current-independent blank, and implemented a blank correction for the method.

HGIS cathodes produce measurable ion current, even with no gas flow. This current was initially \> 10 $\mu$A and dropped exponentially, reaching a baseline of ~100 nA after 5-10 minutes. For the range of currents in this study (3--9 $\mu$A), this baseline current represents a 1--6% contribution to the total measured sample current at the F^14^C of the target current baseline. This can be interpreted as a constant current dependent blank, analogous to a mass-dependent "constant contaminant" blank and may be modeled with the same tools [@haghipour_compound-specific_2019; @hanke_comprehensive_2017; @roberts_14c_2019; @sun_14_2020]. To assess blank contribution and ratio dependence on sample current, vials of live and dead standard gas were measured as the displacement helium flow diluted the samples, producing diminishing current over time. We used these data to develop a current-balance model following methods used to assess mass-dependent or constant contamination blanks in graphite targets [@shah_walter_ultra-small_2015; @sun_14_2020]. The constant contamination blank was also assessed directly by measuring pure helium in vials using the same method. Results from both tests are shown in Figure 3a. The inverse of the sample current is used to linearize the data, allowing assessment with a simple linear model. Fits from modern and dead dilution tests should trend towards the fraction modern of the blank, crossing at the inverse of blank current. Based on this assessment, we estimate the F^14^C of the blank as 0.43 with a current of 110 nA.

```{r fig3, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig3}Assessment of blanks. (a) Current dependent blank model and data. $F^{14}C$ and current of measurements from dilution tests of modern and dead tank standards are used to generate linear fits which should cross at the $F^{14}C$ and current of the constant current blank. Direct determination of $F^{14}C$ and current of the blank using clean targets and helium without $CO_2$ are also shown. (b) Fraction modern of large IAEA C1 and TIRI F carbonate blank samples compared to sample current. Horizontal and vertical bars represent measurement uncertainty in current and ratio, respectively.", echo = FALSE}
knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig3_blanks.pdf"))
```

@@

#options(digits = 3)
options(scipen = 99)
options(decimals = 4)
options(errors.notation = "plus-minus")

@@

Reference materials analyzed with the hybrid gas method were in good agreement with consensus values. Mean difference from consensus for all measurements was F^14^C `r format(cons_sum$fm_diff_mean, digits = 1, nsmall = 4)` shows that the hybrid method has no bias in aggregate. The standard deviation of the difference in F^14^C was `r format(cons_sum$fm_diff_sd, digits = 2, nsmall = 4)`, which may be taken as an estimate of the 1$\sigma$ precision of the method. The mean of propagated measurement errors for these samples was `r format(cons_sum$sig_fm_corr_mean, digits = 2, nsmall = 4)` F^14^C for these samples, which is in good agreement with the method precision above. The method error is about three times larger than the 2--3 ‰ precision of graphite measurements, but is within the range of method precision of other gas sources.

```{r tab1, message = FALSE}
# Summary table
cons_df %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  ungroup() %>% 
  arrange(fm_consensus) %>% 
  mutate (fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd),
          Name = case_when(Name == "TIRI-I" ~ "TIRI I",
                           Name == "C-2" ~ "IAEA C2",
                           Name == "NOSAMS2" ~ "NOSAMS 2")) %>% 
  select(`Sample Name` = Name, 
         `Reference F^14^C` = fm_consensus,
         `HGIS F^14^C ± stdev` = fm_corr_err,
         `Mean per-sample error` = sig_fm_corr_mean,
         N) %>% 
  knitr::kable(digits = c(4, 4, 4, 3, 2),
               caption = "\\label{tab1}Reference material precision and accuracy. Reference 
               values are community consensus values for TIRI I and C2, and the 
               NOSAMS internal graphite value for NOSAMS2. The mean of per-sample 
               error estimates is included for comparison to the reproducibilty 
               of the samples.")

@@

```{r fig4, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig4}Agreement of standards with graphite consensus values. Difference of HGIS measurements from consensus for individual samples are plotted against their expected value. Expected values are spread to avoid overlapping points. Error bars represent 1$\\sigma$ sample uncertainty. The black line is the mean difference from expected value for all measurements, and the blue lines are one standard deviation of all measurements.", echo = FALSE}
knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig4_consensus_samples.pdf"))
```

@@

```{r fig5, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig5}Comparison of Haiti core samples measured as graphite and via HGIS. (a) Difference between HGIS and graphite measurements of the same materials. Difference between HGIS and graphite measurements is shown on the vertical axis and plotted against the mean of the two measurements. Error bars are 1$\\sigma$ measurement uncertainty for each sample. Graphite measurements are included to allow comparison of uncertainty. Bayesian age-depth models produced using HGIS (b) and graphite measurements (c). Grey subplots show probability distribution of calibrated ages. (d) Combined age-depth prediction and 95\\% confidence intervals for HGIS and graphte. (e) Offset of HGIS and graphite values from the mean of the models with confidence intervals.", echo = FALSE}
knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig5_core_carbonates.pdf"))
```

@@

Age models were developed for both the HGIS measurements (Figure 5b) and the traditional graphite measurements (Figure 5c) of the core carbonates. We used the Bchron Bayesian radiocarbon chronology package [@haslett_simple_2008] to develop these models. We applied a correction for the regional marine reservoir effect with a $\Delta$ R of -216 ± 92 years based on the Oxcal model [@bronk_ramsey_bayesian_2009] with 10 reservoir points. Samples with pre-bomb radiocarbon ages were calibrated using the Marine20 curve @heaton_marine20marine_2020. The NH zone 2 curve from Hua et. al @hua_atmospheric_2013 was used for calibration of postmodern dates, but these measurements were not included in the age model due to uncertainties in reservoir correction. The age-depth curves produced by the two datasets agree within the uncertainty of the chronologies (Figure 5d & 5e), with the largest differences being in the uncertainty of the individual dates.

@@

We developed a method for the analysis of carbonates using an open-split interface to a hybrid sputter gas ion source. This system produces results that agree with graphite measurements within the approximately 1% method error from 8--20 mg of carbonate. Sample preparation takes about 10 min. of person time and 6 h of automated processing and reaction time, and 20 min. of measurement time per sample. The simplicity of of this approach can potentially reduce sample cost by eliminating graphitization, and may improve reliability and reduce maintenance compared to more complex sample inlet designs.
